Unsupervised embedding of trajectories captures the latent structure of mobility

### 2021-01-22 paper/current/main.tex
edits to methods new math

The negative sampling is a biased estimator~\autocite{Chia2010,Dyer2014} that optimizes a model different from the {\it word2vec}.
What is the different model?

The negative sampling trains the word2vec by a binary classification task as follows.
One samples a center-context word pair $(i,j)$ from the given data as a positive pair, and sample $k$ negative pairs $(i,\ell)$ from noise distribution $p_0(\ell)=P^\gamma(\ell)$. (Remind that $P(j)$ is the
fraction of $j$ in the given data, and $\gamma>0$ is a parameter).
Then, given a center-context word pair $(i,j)$, one predicts whether it is a positive or negative sample using sigmoid function:
\begin{align}
	\label{eq:sigmoid}
	\frac{1}{1 + \exp\left( - \bm{u}_j \cdot \bm{v}_i \right)}.
\end{align}

The negative sampling does not guarantee that the embedding converges to the optimal one, even if one increases the number of training samples and iterations~\autocite{Chia2010,Dyer2014}.
This inconsistency problem is addressed by the noise contrastive estimation (NCE) \autocite{Chia2010}, which fits a probability model in form:
\begin{align}
	\label{eq:nce}
	f(x) / \sum_{x'} f(x'),
\end{align}
where $f(x)$ is the unnormalized probability for an instance $x$.
As is the case for the negative sampling, the NCE fits the model based on a binary classification task, with positive and negative samples randomly drawn from the given data and a noise distribution $p_0$, respectively, using a bias-corrected sigmoid function, \ie
\begin{align}
	\label{eq:sigmoid2}
	\frac{1}{1 + \exp\left( -f(x) + \log p_0(x) \right)}.
\end{align}
In case of the word2vec, the instance is the dot similarity $x=\bm{u}_i \cdot \bm{v}_j$,  $f(x)=\exp(x)$, and $p_0(j)=P^\gamma(j)$.

>>>>

Negative sampling is a biased estimator~\autocite{Chia2010,Dyer2014} that optimizes a distinct model  from {\it word2vec}.
What is the different model?

Negative sampling trains {\oit word2vec} using a binary classification task as follows.
A center-context word pair $(i,j)$ is sampled from the given data as a positive pair, and $k$ negative pairs $(i,\ell)$ are sampled from a noise distribution $p_0(\ell)=P^\gamma(\ell)$. ($P(j)$ is the fraction of $j$ in the data, and $\gamma>0$ is a parameter).
Then, given a center-context word pair $(i,j)$, the model predicts whether it is a positive or negative sample using sigmoid function:
\begin{align}
	\label{eq:sigmoid}
	\frac{1}{1 + \exp\left( - \bm{u}_j \cdot \bm{v}_i \right)}.
\end{align}

This procedure does not guarantee that the embedding optimally converges, even when increasing the training samples and iterations~\autocite{Chia 2010,Dyer2014}.
This problem is addressed by the noise contrastive estimation (NCE) \autocite{Chia2010}, which fits a probability model in the form:
\begin{align}
	\label{eq:nce}
	f(x) / \sum_{x'} f(x'),
\end{align}
where $f(x)$ is the unnormalized probability for an instance $x$.
As in negative sampling, NCE fits the model based on a binary classification task, with positive and negative samples randomly drawn from the given data and a noise distribution $p_0$, respectively, using a bias-corrected sigmoid function:
\begin{align}
	\label{eq:sigmoid2}
	\frac{1}{1 + \exp\left( -f(x) + \log p_0(x) \right)}.
\end{align}
In case of the {\it word2vec}, the instance is the dot similarity $x=\bm{u}_i \cdot \bm{v}_j$,  $f(x)=\exp(x)$, and $p_0(j)=P^\gamma(j)$.
